---
title: "Which type of review should I do?"
author: "Matt"
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

There are a number of decision trees to aid choosing a review methodology. Here I have adapted the ["Knowledge Synthesis Decision Tool" from Library Services, University Health Toronto](https://guides.hsict.library.utoronto.ca/SMH/systematic/decisiontool). I removed the decision point around narrative review as, although an [argument](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6001568/) is put forward for them I can not see that the risk of bias for decision makers can be adequately reduced to make these useful.  



```{r, echo=FALSE,layout="l-screen-inset"}
library(DiagrammeR)
grViz(diagram = "digraph flowchart {
      # define node aesthetics
      node [fontname = Arial, shape = rect, color = DeepSkyBlue, style = filled, fontcolor = White]        
      tab1 [label = 'Do you have >3 on the review team?']
      tab2 [label = 'Do you have 9 to 18 months to conduct the review?']
      tab3 [label = 'Reviews need a multi-person team for unbiased article screening but see Rapid Reviews']
      tab4 [label = 'Do you have a broad topic or multiple questions?']
      tab5 [label = 'Are you an expert or know the literature on this topic well?']
      tab6 [label = 'Scoping Review/Systematic Map', color=DarkGreen]
      tab7 [label = 'Are there multiple published Systematic Reviews on your topic?']
      tab8 [label = 'Umbrella Review or Review of Reviews',color=DarkGreen]
      tab9 [label = 'Do you have a well formatted research question,e.g. PICO and inclusion/exclusion criteria?']
      tab10 [label = 'Do you think the included studies are homogenous enough to statistically evaluate, synthesise and summarise results objectively?']
      tab11 [label = 'A well formulated question is vital - speak to your library search specialist']
      tab12 [label = 'Systematic Review & meta-analysis',color=DarkGreen]
      tab13 [label = 'Systematic Review',color=DarkGreen]
      tab14 [label='Rapid Review',color=DarkGreen]
      tab15 [label='Add subject experts to join the team and use a Rapid Review']
# set up node layout
      tab1 -> tab2[label=Yes];
      tab1 -> tab3[label=No];
      tab2 -> tab4[label=Yes];
      tab2 -> tab5[label=No];
      tab5 -> tab14[label=Yes];
      tab5 -> tab15[label=No];
      tab4 -> tab6[label=Yes];
      tab4 -> tab7[label=No];
      tab7 -> tab8[label=Yes];
      tab7 -> tab9[label=No];
      tab9 -> tab10[label=Yes];
      tab9 -> tab11[label=No];
      tab10 -> tab12[label=Yes];
      tab10 -> tab13[label=No];
      tab11 -> tab9[label='Rethink the question']
      tab3 -> tab14
      }

      ")


```


## For further information on these review types:
* [Rapid review](https://drmattg.github.io/Evidence_synthesis/posts/What_is_rapid_review/)

* [Scoping Reviews/Systematic Map](https://drmattg.github.io/Evidence_synthesis/posts/What_is_systematic_mapping/)

* [Umbrella Review/Review of reviews](https://drmattg.github.io/Evidence_synthesis/posts/What_is_review_review/)

* [Systematic Review](https://drmattg.github.io/Evidence_synthesis/posts/What_is_systematic_review/)

* [Systematic Review & meta-analysis](https://drmattg.github.io/Evidence_synthesis/posts/What_is_meta_analysis/)